---
title: "Exam 3"
author: "McKay Jones"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Load and Clean Data

In part 1, I'm going to load and clean the FacultySalaries_1995.csv file and re-create a graph.

First, let's load some required projects.
```{r include=FALSE}
library(tidyverse)
library(modelr)
library(easystats)
library(GGally)
library(caret)
library(broom)
```



I will load the data using the read.csv command, and save it to a 'dat' object.

```{r}
dat <- read.csv('FacultySalaries_1995.csv')
```


Let's take a look at this data so we know what to clean.

```{r}
glimpse(dat)
```

I'm cleaning the data by removing any 'VIIB' schools, and using the pivot_longer command to combine all of the professors' positions and salaries into two rows. This will make it easier to construct Fig1.

In the newly constructed professor_type column, I made sure to remove redundant information ('Avg' in the beginning of the string and 'ProfSalary' at the end of the string). Now, all of the values in the professor_type column are 'Assist', "Assoc', and 'Full' (instead of 'AvgAssistProfSalary', for example).

This 'tidied' data was saved to a dat_clean object for dowmstream analyses.

```{r}
dat_clean <- dat %>%
  filter(Tier != 'VIIB') %>%
  pivot_longer(ends_with('Salary'), names_to = 'professor_type',
               names_prefix = "Avg",
               values_to = 'avg_salary') %>%
  mutate(professor_type = sub("ProfSalary", "", professor_type))
```

I then constructed the desired boxplot image, using facet_wrap() to group all the salaries by school rank. It looks like professors at Rank I schools tend to have higher salaries.

```{r}
img_1 <- dat_clean %>%
  ggplot(aes(x = professor_type,
             y = avg_salary,
             fill = professor_type)) +
  geom_boxplot() +
  facet_wrap(~Tier) +
  theme_minimal() +
  labs(x = 'Rank', y = "Salary") +
  guides(fill = guide_legend(title = "Rank")) +
  theme(axis.text.x = element_text(angle = 45))
```


Let's view the image and save it.

```{r}
img_1

ggsave('Fig1.png', plot = img_1, width = 3, height = 5)

```

## Part 2: ANOVA Model

I am building an ANOVA model that tests the influence of State, School Tier, and Professor Rank/Type on "Salary" with NO interactions between the predictors.

```{r}
mod_anova <- aov(data = dat_clean, avg_salary~State+Tier+professor_type)

summary(mod_anova)
```


## Part 3: Tidy Juniper Dataset

This data set originated from a collaboration between Young Living Inc. and UVU Microbiology.

First, I will load this data into a jun_dat object and take a look.

```{r}
jun_dat <- read.csv("Juniper_Oils.csv")

glimpse(jun_dat)
```
Let's tidy this dataset. First, I saved the compounds I want to analyze into a 'chems' vector. To match the names in this vector, I replaced all of the periods in the original column names with '-'.

For 'compound 1' and 'compound 2', which are separated by a space, I saved these two rows into another object with their '-' replaced with spaces. Then, I used bind_cols to add them back to the original data frame.

```{r}
chems <- c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene","alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene","beta-chamigrene","cuparene","compound 1","alpha-chamigrene","widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol","beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol","compound 2","thujopsenal")


jun_dat2 <- jun_dat %>%
  rename_with(~ gsub("\\.", "-", .x))

compounds <- jun_dat2 %>%
  select(starts_with('compound')) %>%
  rename_with(~ gsub("\\-", " ", .x))

jun_dat3 <- jun_dat2 %>%
  select(-starts_with('compound')) %>%
  bind_cols(compounds)
```

Now, I will use the pivot_longer command on the chemicals and Mass-Spec concentrations of each one. This will be helpful for the figure we want to make.

```{r}
jun_dat4 <- jun_dat3 %>%
  pivot_longer(chems, names_to = 'chemical',
               values_to = 'concentration')
```

## Part 4: Make a Graph

I will make a graph, with x = YearSinceBurn, y = Concentration, and facet = ChemicalID

```{r}
img_2 <- jun_dat4 %>%
  ggplot(aes(x = YearsSinceBurn,
             y = concentration)) +
  geom_smooth() +
  facet_wrap(~chemical, scales = 'free_y')

img_2
```

## Part 5: Find chemicals that are significantly affected

Now, I am going to use a generalized linear model to find which chemicals show concentrations that are significantly affected by "Years Since Burn".

```{r}
models <- jun_dat4 %>%
  group_by(chemical) %>%
  do(tidy(glm(concentration ~ YearsSinceBurn, data = .)))

sig_chems <- models %>%
  filter(term == "YearsSinceBurn",
         p.value < 0.05) %>%
  arrange(p.value)

sig_chems
```

